Compositional equivalences based on Open pNets

### 2020-02-21 2016-Weak-FH-Bisim/WeakBisim.tex

    For readability, we did not include the full details of the Implementation weak open transitions in Figure \ref{SimpleProtCounter:ImplOA}, they will appear below. In this automaton, many WOT are similar, and numbered accordingly as 3, 3', 3'', and 8, 8', 8'' respectively: they only differ by the name of the state variables in their respective source or target states.
\end{example}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=9cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{SimpleProtCounter:ImplOA}
\end{figure}

Now let us give some details about the construction of the weak automaton of the implementation pNet, obtained by application of the weak rules as explained above. Here we denote as $post_n$ the effect (= substitution) of the strong open transition $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

$post_1 = post_2 = ()$

@@

$post_7 = (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_ec}\gets \texttt{m1\_ec})$

$post_8 = post_9 = ()$
\medskip

@@

We can now show some of the weak OTs of Figure \ref{SimpleProtCounter:SpecOA-partial} (the full table is included in Appendix \ref{Appendix:FullExample}) :
\smallskip

$ 3' = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{m1\_msg}\gets m, \texttt{m1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
  \ \forall n\ge 0$

         $ 4 = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

$ 8 = \openrule
         {\{\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(m,c)}}\subst{(m\gets \texttt{r1\_msg}, c\gets \texttt{r1\_ec}}} 202}
         $

@@

%% \TODO{move text below}

%% \begin{property}[Decomposition of weak open transitions] \TODO{TODO}
%% Let $A$ be the open automaton that is the semantics of a \pNet $P$, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
%% Suppose we have the following weak open transitions in $\WT$:
%% \[\openrule
%%          {
%%            \set{\gamma_1},\Pred_1,\Post_1   }
%%          {s \OTWeakarrow {\tau} s_1}
%% \qquad 
%% \openrule
%%          {
%%            \set{\gamma_2},\Pred_2,\Post_2   }
%%          {s_1 \OTWeakarrow {\alpha} s_2}
%% \qquad
%% \openrule
%%          {
%%            \set{\gamma_3},\Pred_3,\Post_3   }
%%          {s_2 \OTWeakarrow {\tau} s_3}
%% \]
%% Then there is ag weak  open transition of the following form in $\WT$:
%% \[\openrule
%%          {
%%            \set{\gamma},\Pred,\Post   }
%%          {s \OTWeakarrow {\alpha} s_3}
%% \]
%% Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
%% \[
%% \begin{array}{rcl}
%% \gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
%% \Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
%% \Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
%% \end{array}\]
%% \end{property}

@@

\subsection{Running example}
- In section \ref{def:weakOT} we have shown the full satured weak automaton for both the specification and the implementation of the simple protocol.

\TODO{Je ne pense pas que nous ayons un algo de bisim weak operationnel pour la publi, mais ca n'est pas completement impossible nonp[l;us...  Au pire, argumenter intuitivement sur papier, en donnat les classes d'equivalence et les predicats.}

  \begin{itemize}
    \item[] Both pNets trivialy verify the ``non-observability''
      condition: the only vectors having $\tau$ as an action of a
      sub-net are of the form ``$< -, \tau, -> -> \tau$''.
  \end{itemize}

  Now consider the following triples:

@@

$ 3' = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{m1\_msg}\gets m, \texttt{m1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
  \ \forall n\ge 0$

$ 3'' = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True, 
    (\texttt{m2\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 220}
  \ \forall n\ge 0$

         $ 4 = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

@@

$ 8 = \openrule
         {\{\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(m,c)}}\subst{(m\gets \texttt{r1\_msg}, c\gets \texttt{r1\_ec}}} 202}
         $

\end{document}

>>>>

    \TODO{introduce the notation for parametrized OTs}

    For readability, we did not include the full details of the Implementation weak open transitions in Figure \ref{SimpleProtCounter:ImplOA}, they will appear below.
    First, we the weak OT loops on state $ostate{000}$ are also present in all other states, we did not repeat them. Then many WOTs are similar, and numbered accordingly as 3, 3', 3'', and 8, 8', 8'',8''' respectively: they only differ by the name of the state variables in their respective source or target states.
\end{example}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=11cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{SimpleProtCounter:ImplOA}
\end{figure}

Now let us give some details about the construction of the weak automaton of the implementation pNet, obtained by application of the weak rules as explained above. We concentrate on transitions $WT_3$ and $WT_4$. Let us denote as $post_n$ the effect (= substitution) of the strong open transitions $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

@@

@@

We can now show some of the weak OTs of Figure \ref{SimpleProtCounter:ImplOA} (the full table is included in Appendix \ref{Appendix:FullExample}) :
\smallskip

$ WI_1 = \openrule
{\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \ \forall n\ge 0$

  $ WI_4(n) = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

$ WI_8(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(r1-msg,r1-ec +n)}\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(r1-msg,r1-ec +n)}}} 202}
         \ \forall n\ge 0$

@@

@@

\subsection{Running example}
In section \ref{def:weakOT} we have shown the full satured weak automaton for both the specification and the implementation of the simple protocol. We will show here how we can check if some given relation between these two automata is a weak FH-Bisimulation.

\medskip
\noindent
Preliminary remarks:
  \begin{itemize}
    \item Both pNets trivialy verify the ``non-observability''
      condition: the only vectors having $\tau$ as an action of a
      sub-net are of the form ``$< -, \tau, -> -> \tau$''.
      \item We must take care of variable name conflicts: in our example, the state variables of the 2 systems already have different names, but the action parameters occuring in the transitions (m, msg, ec) are the same, that is not correct. In the tools, this is managed by the static semantic layer; in the following example, we have renamed all conflicting variables with 1 for the Spec, and 2 for the Impl.
  \end{itemize}

  Now consider relation $\mathcal{R}$ defined by the following triples:

  \begin{tabular}{|c|c|l|}
\hline
    Spec state & Impl state & Predicate\\
    \hline
    b0 & $\ostate{000}$ & True\\
    b0 & $\ostate{202}$ & True\\
    b1 & $\ostate{100}$ & $\texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec}$\\
    b1 & $\ostate{210}$ & $\texttt{b1\_msg = m1\_msg} \land \texttt{b1\_ec = m1\_ec}$\\
    b1 & $\ostate{220}$ & $\texttt{b1\_msg = m2\_msg} \land \texttt{b1\_ec = m2\_ec}$\\
    b1 & $\ostate{201}$ & $\texttt{b1\_msg = r1\_msg} \land \texttt{b1\_ec = r1\_ec}$\\
    \hline
    \end{tabular}

  \medskip
  Checking that $\mathcal{R}$ is a weak FH-Bisimulation means checking, for each of these triples, that each WOT of one the states corresponds to a set of WOTs of the other, using the conditions from Definition \ref{def-Weak-bisim}.
  We give here two examples.

  Consider the first triple from the table, and transition $WS_3(n)$ from state $\ostate{b0}$ for some fixed integer $n$. Its easy to guess that it will correspond to $WI_3(n)$ of $\ostate{000}$.

  $ WS_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{b1\_msg}\gets \texttt{m}, \texttt{b1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  $

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  $

  Let check formally the conditions:

\begin{itemize}
  \item Their sets of active (non-silent) holes is the same: $\{P\}$.
  \item Triple ($\ostate{b1},  \ostate{100},  \texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec}$) is in $\mathcal{R}$.
  \item The verification condition \\
    $\Pred \land \Pred_{OT}%\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}\!
   \left( \forall j\in J_x. \vis{\beta_j}\!=\!\gamma_{jx}\! \land\! \Pred_{OT_x}
     \!\land\! \alpha\!=\!\alpha_x\! \land\!  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)}$

\medskip Gives us:

\TODO{Insert here $\forall \phi(m1,m2)$ }

$True \land True\hspace{1cm} \implies\\ 
\texttt{p-send(m1)} = \texttt{p-send(m2)}
\land True
\land \underline{\texttt{in(m1)}} = \underline{\texttt{in(m2)}}
\land (\texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec})
\subst{(\texttt{b1\_msg}\gets \texttt{m}, \texttt{b1\_ec}\gets n)
  \uplus (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
$

\medskip That reduces to:

$\forall \texttt{m1}. \exists \texttt{m2}.
(\texttt{p-send(m1)} = \texttt{p-send(m2)}
\land \texttt{in(m1)} = \texttt{in(m2)}
\land \texttt{m}=\texttt{m} \land n=n)$

\medskip That is a tautology.

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \ \forall n\ge 0$

$ WI_4(n) = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

@@

$ WI_8(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(r1-msg,r1-ec +n)}\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(r1-msg,r1-ec +n)}}} 202}
         \ \forall n\ge 0$

$ WI_9(n) = \openrule
         {\{\}, True, (\texttt{m1\_ec}\gets \texttt{m1\_ec}+n)}
         {210 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$
\end{document}
